\chapter{Conclusions}

The goal for this PhD thesis is to incorporate techniques of task and thread scheduling in order to fully utilize asymmetric systems.
The main contributions of the thesis rely on the efficient exploitation of future asymmetric multi-core systems in terms of performance and energy efficiency as well as on the development of future asymmetric systems that fit the needs of high performance computing.
Existing parallel scientific applications will become portable when moving from a traditional multi-core to an asymmetric multi-core system. 
Our useful observations throughout this study will also contribute and give guidelines for the design of the future multi-core asymmetric systems for high performance computing.

%goals are performance and energy efficiency as well as the portability of existing applications from the traditional homogeneous multi-cores to the new asymmetric multi-core systems.

%From our current results we have seen that current asymmetric multi-core systems are not ready to efficiently run out of the box high performance scientific applications and that the most efficient way is by using a task-based approach.
%Moreover, we have seen the potential of the criticality-aware task schedulers to speed up dependency-intensive applications and take advantage of the asymmetric compute resources is very high but sometimes comes at the cost of high additional overhead.
%We expect that adding one more scheduling layer will help eliminate the scheduling overheads of the smart heterogeneous scheduling approaches and boost performance and energy efficiency of such architectures.

Our current results have shown that the state-of-the-art asymmetric multi-core systems are not ready to efficiently run out-of-the-box high performance applications and that the most efficient way is by using a task-based approach.
This increases the need of research in this direction through the paths of scheduling and thread migration as described in the previous Chapters.
In our first attempts to follow these paths, we have seen the high potential of the criticality-aware task schedulers to speed up dependency-intensive applications and take advantage of the asymmetric compute resources.

We are optimistic that following our second research approach of runtime thread migration will also contribute positively.
The greatest challenge will be to increase performance without sacrificing energy, thus the dynamic search for the appropriate assistant core for the runtime thread has to consider all these obstacles.
We expect that this approach will also influence designers to consider the use of assistant cores in the future asymmetric multi-cores.
Finally, in our last and most complete approach we will need to synchronize all of our tools (e.g. scheduling and thread migration) to adapt to the runtime circumstances and boost performance with decent energy consumption.

Since a part of this work is already complete, we expect that our goals will be successfully accomplished and this study will be a useful reference for the future research.



%Our approach on dynamic runtime thread migration will further improve performance and energy efficiency and we expect that it will 



%The goal for this PhD thesis is to incorporate techniques from approximate computing to improve the performance in the scientific computing domain without incurring too much energy consumption overhead or drastically altering the current parallel programming paradigm. 
%~\\ \\
%It is a slight paradigm shifting from the traditional scientific computing ideology: to execute applications in a very-high-precision fashion. By loosing some of the precision 
%restrictions at some points of the execution one is able to open up more parallelism to boost the performance with the help of the runtime system yet maintaining the quality of the final
%results.
%~\\ \\
%As a novel approach, we can see the obstacle that lies beyond. Identifying the degree of approximation, realizing the right type of approximation, the applicability of the techniques 
%etc. all are still big questions. Yet with a right mindset and the progress we already have we are confident towards this approach.